---
title: "Life Expectancy"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results='asis', echo=FALSE}
cat('<link rel="stylesheet" type="text/css" href="css/life.css" />')
```

<script src="https://code.jquery.com/jquery-3.1.1.min.js"></script>
<script type="text/javascript" src="js/d3.js"></script>
<script type="text/javascript" src="js/d3.time.js"></script>
<script type="text/javascript" src="js/d3.csv.js"></script>

```{js, echo=FALSE}
      var _gaq = _gaq || [];
      _gaq.push(['_setAccount', 'UA-865100-8']);
      _gaq.push(['_trackPageview']);

      (function() {
        var ga = document.createElement('script'); ga.type = 'text/javascript'; ga.async = true;
        ga.src = ('https:' == document.location.protocol ? 'https://ssl' : 'http://www') + '.google-analytics.com/ga.js';
        var s = document.getElementsByTagName('script')[0]; s.parentNode.insertBefore(ga, s);
      })();

```

People are living longer around the world, some more so than others. Select a region (as defined by World Bank) below to compare, or roll over to the graph to highlight countries. Copied from  [flowingdata.com](http://projects.flowingdata.com/life-expectancy/).

<div class="btn-group btn-group-justified" id="filters">
<a id="EAS" class="btn btn-default">East Asia &amp; Pacific</a>
<a id="SAS" class="btn btn-default">South Asia</a>
<a id="ECS" class="btn btn-default">Europe &amp; Central Asia</a>
<a id="MEA" class="btn btn-default">Middle East &amp; North Africa</a>
<a id="SSF" class="btn btn-default">Sub-Saharan Africa</a>
<a id="LCN" class="btn btn-default">Latin America &amp; Caribbean</a>
<a id="NAC" class="btn btn-default">North America</a>
</div>

<div id="blurb">
<div id="default-blurb">
## World

The average life expectancy in the world in 2009 was 69 years.
</div>
<div id="blurb-content"></div>
</div>
<div id="vis"></div>

```{js}

//Get how wide our page is in pixels so we can draw our plot in it
var page_width = $("#vis").width();
  
// set the dimensions and margins of the graph
var margin = 40,
    width = page_width - 2*margin,
    height = page_width*0.8 - 2*margin;

var regions = { "SAS": "South Asia" , "ECS": "Europe and Central Asia", "MEA": "Middle East & North Africa", "SSF": "Sub-Saharan Africa", "LCN": "Latin America & Caribbean", "EAS": "East Asia &amp; Pacific", "NAC": "North America" },
	w = width,
	h = height,
	margin = 30,
	startYear = 1960, 
	endYear = 2010,
	startAge = 20,
	endAge = 80,
	y = d3.scale.linear().domain([endAge, startAge]).range([0 + margin, h - margin]),
	x = d3.scale.linear().domain([1960, 2009]).range([0 + margin -5, w]),
	years = d3.range(startYear, endYear);

var vis = d3.select("#vis")
    .append("svg:svg")
    .attr("width", w)
    .attr("height", h)
    .append("svg:g")
    // .attr("transform", "translate(0, 600)");

			
var line = d3.svg.line()
    .x(function(d,i) { return x(d.x); })
    .y(function(d) { return y(d.y); });
					

// Regions
var countries_regions = {};
d3.text('data/country-regions.csv', 'text/csv', function(text) {
    var regions = d3.csv.parseRows(text);
    for (i=1; i < regions.length; i++) {
        countries_regions[regions[i][0]] = regions[i][1];
    }
});

var startEnd = {},
    countryCodes = {};
d3.text('data/life-expectancy-cleaned-all.csv', 'text/csv', function(text) {
    var countries = d3.csv.parseRows(text);
    
    for (i=1; i < countries.length; i++) {
        var values = countries[i].slice(2, countries[i.length-1]);
        var currData = [];
        countryCodes[countries[i][1]] = countries[i][0];
        
        var started = false;
        for (j=0; j < values.length; j++) {
            if (values[j] != '') {
                currData.push({ x: years[j], y: values[j] });
            
                if (!started) {
                    startEnd[countries[i][1]] = { 'startYear':years[j], 'startVal':values[j] };
                    started = true;
                } else if (j == values.length-1) {
                    startEnd[countries[i][1]]['endYear'] = years[j];
                    startEnd[countries[i][1]]['endVal'] = values[j];
                }
                
            }
        }
        vis.append("svg:path")
            .data([currData])
            .attr("country", countries[i][1])
            .attr("class", countries_regions[countries[i][1]])
            .attr("d", line)
            .on("mouseover", onmouseover)
            .on("mouseout", onmouseout);
    }
});  
    
vis.append("svg:line")
    .attr("x1", x(1960))
    .attr("y1", y(startAge))
    .attr("x2", x(2009))
    .attr("y2", y(startAge))
    .attr("class", "axis")

vis.append("svg:line")
    .attr("x1", x(startYear))
    .attr("y1", y(startAge))
    .attr("x2", x(startYear))
    .attr("y2", y(endAge))
    .attr("class", "axis")
			
vis.selectAll(".xLabel")
    .data(x.ticks(5))
    .enter().append("svg:text")
    .attr("class", "xLabel")
    .text(String)
    .attr("x", function(d) { return x(d) })
    .attr("y", h-10)
    .attr("text-anchor", "middle")

vis.selectAll(".yLabel")
    .data(y.ticks(4))
    .enter().append("svg:text")
    .attr("class", "yLabel")
    .text(String)
	.attr("x", 0)
	.attr("y", function(d) { return y(d) })
	.attr("text-anchor", "right")
	.attr("dy", 3)
			
vis.selectAll(".xTicks")
    .data(x.ticks(5))
    .enter().append("svg:line")
    .attr("class", "xTicks")
    .attr("x1", function(d) { return x(d); })
    .attr("y1", y(startAge))
    .attr("x2", function(d) { return x(d); })
    .attr("y2", y(startAge)+7)
	
vis.selectAll(".yTicks")
    .data(y.ticks(4))
    .enter().append("svg:line")
    .attr("class", "yTicks")
    .attr("y1", function(d) { return y(d); })
    .attr("x1", x(1959.5))
    .attr("y2", function(d) { return y(d); })
    .attr("x2", x(1960))

function onclick(d, i) {
    var currClass = d3.select(this).attr("class");
    if (d3.select(this).classed('selected')) {
        d3.select(this).attr("class", currClass.substring(0, currClass.length-9));
    } else {
        d3.select(this).classed('selected', true);
    }
}

function onmouseover(d, i) {
    var currClass = d3.select(this).attr("class");
    d3.select(this)
        .attr("class", currClass + " current");
    
    var countryCode = $(this).attr("country");
    var countryVals = startEnd[countryCode];
    var percentChange = 100 * (countryVals['endVal'] - countryVals['startVal']) / countryVals['startVal'];
    
    var blurb = '<h2>' + countryCodes[countryCode] + '</h2>';
    blurb += "<p>On average: a life expectancy of " + Math.round(countryVals['startVal']) + " years in " + countryVals['startYear'] + " and " + Math.round(countryVals['endVal']) + " years in " + countryVals['endYear'] + ", ";
    if (percentChange >= 0) {
        blurb += "an increase of " + Math.round(percentChange) + " percent."
    } else {
        blurb += "a decrease of " + -1 * Math.round(percentChange) + " percent."
    }
    blurb += "</p>";
    
    $("#default-blurb").hide();
    $("#blurb-content").html(blurb);
}
function onmouseout(d, i) {
    var currClass = d3.select(this).attr("class");
    var prevClass = currClass.substring(0, currClass.length-8);
    d3.select(this)
        .attr("class", prevClass);
    // $("#blurb").text("hi again");
    $("#default-blurb").show();
    $("#blurb-content").html('');
}

function showRegion(regionCode) {
    var countries = d3.selectAll("path."+regionCode);
    if (countries.classed('highlight')) {
        countries.attr("class", regionCode);
    } else {
        countries.classed('highlight', true);
    }
}

```

```{js}
$(document).ready(function() {
    $('#filters a').click(function() {
        var countryId = $(this).attr("id");
        $(this).toggleClass(countryId);
        showRegion(countryId);
    });
    
});
```

<!-- part of the requirement for the javascipt to run -->